Copulas and their potential for ecology

### 2019-12-13 Paper.Rmd
abstract word count:253
new responses file and working on some changes in response to ref 1

\noindent \emph{The number of words in the abstract:} 263

@@

The datasets we used included environmental, species-trait, phenological, 
population, community, and ecosystem functioning data (Table \ref{tab_data_info}), selected to span multiple fields and levels
of organization within ecology. Copula structure describing 
the relationship between atmospheric weather variables such as rainfall or wind speed, 
as measured in multiple locations through time, has been 
examined previously in the meteorological literature [e.g., @Serinaldi2008; @li2013]. We therefore 
examined environmental variables from the soil instead, using the Rapid Carbon 
Assessment database [RaCA; @Wills2014].
<!--***DAN: Add Wills and Loecke pending reference if published in time--> 
The database comprises measurements of soil organic carbon and total soil 
nitrogen (Mg C or N per hectare of soil surface) from `r nrow(d_soilCN)` 
locations across the coterminous United States (Fig. \ref{SM-fig_SoilCNmapUSA} for sampling locations).
Species-trait data were average species basal metabolic rate (BMR, KJ per hour) and 
body mass (grams) for `r nrow(d_birdsBMR)`  species of birds [@Mcnab_2009] 
and `r nrow(d_mammalsBMR)`  species of mammals [@Mcnab_2008].
These data have been much studied, but to our knowledge the copula structure 
of the dependence between metabolic rate and body mass has not been examined. 
Species-trait data such as these reflect the coevolution of the two traits.
Phenological data were aphid first-flight dates from 10 locations (Fig. \ref{SM-fig_aphid_map}) 
across the United Kingdom (UK) for 20 aphid species (Table \ref{SM-tab_aphid_info}), for the 
35 years 1976 to 2010. These time series were computed from the Rothamsted Insect 
Survey (RIS) suction-trap dataset [@Harrington2014; @Bell2015]. The first of our two 
population-level datasets was also derived from the RIS suction-trap data, and comprised 
total counts of aphids trapped for the same locations, species, and years. 
Our second population dataset comprised average annual plankton abundance estimates 
for 14 locations (Fig. \ref{SM-fig_plankton_map})
in the North Sea and British seas for 22 taxa (Table \ref{SM-tab_plankton_info}) for the 
56 years 1958 to 2013. The locations are $2^\circ$ by $2^\circ$ grid cells.
These data were computed from the Continuous Plankton Recorder survey hosted by the 
UK Marine Biological Survey.
Community-level data, obtained from the Cedar Creek Ecosystem Science Reserve, were 
plant aboveground biomass [@ccdataple120] and Shannon's diversity index (computed from plant 
species percent cover data [@ccdatapce120]) for `r nrow(d_cc[[5]])` plots [@ccdataplotse120], 
each 9m by 9m, as sampled in the years 1996-2000 and 2007, each year 
analyzed separately [@Tilman2001; @Tilman2006]. 
Finally, ecosystem functioning data were methane (CH$_{4}$) fluxes between the soil or 
water surface and the troposphere, measured at 13 locations at daily to weekly intervals 
from September 2015 to September 2016 at the Great Miami Wetland mitigation bank in Trotwood, 
Ohio [@Holland1999; @Jarecke2016; @smyth2019using]. Each included location was measured on at least 50 dates. See Appendix \ref{SM-Data_details} for additional data details.

@@

We fit 16 bivariate copula families, exhibiting a variety of 
lower- and upper-tail dependence characteristics, with bivariate datasets. The purpose
of using a large collection of families was solely to include a 
variety of alternative dependence structures to have a robust model
selection and multi-model inference procedure. For that purpose, it 
is not important for the reader to
understand the details of these copulas,
and, additonally, these copulas have been 
described in more detail elsewhere [@Joe1997]; so 
we only briefly identify each family, say a few words about tail dependence 
of the families, and provide pictorial descriptions in the Appendices.
We used several families that can exhibit positive lower-tail dependence 
(of strength depending on parameters) and 
zero upper-tail dependence: the Clayton, survival Gumbel, survival Joe
and survival BB6 families. We used several families that can exhibit positive
upper-tail dependence (strength depends on parameters) and zero lower-tail
dependence: the survival Clayton, Gumbel, Joe and BB6 families. We used
families that show zero upper- and lower-tail dependence: the normal
and Frank families. These families have pdfs symmetric about the line
$v=-u+1$. We also used several families that can show both upper- and
lower-tail dependence, in relative amounts that depend on parameter
values: the BB1, survival BB1, BB7 and survival BB7 families. 
We also used the BB8 family, which shows zero lower-tail dependence,
and zero upper-tail dependence except for a boundary case for the parameters.
And we used the survival BB8 copula, which shows zero upper-tail dependence, 
and zero lower-tail dependence except for a boundary case.
"Survival" families are rotations of the copula with the similar name by
180 degrees. 
See @Joe1997 for details on all these families. We used the implementations
of these families provided in the `VineCopula` package for the R programming
language. See Figs \ref{figped_N}-\ref{figped_BB1} and \ref{SM-figped_SC} and \ref{SM-figped_G}-\ref{SM-figped_SBB8} 
for visual depictions of the pdfs of these copulas and how
pdfs and tail dependence are influenced by the parameters.

>>>>

\noindent \emph{The number of words in the abstract:} 253

@@

The datasets we used included environmental, species-trait, phenological, 
population, community, and ecosystem functioning data (Table \ref{tab_data_info}), selected to span multiple fields and levels
of organization within ecology. Copula structure describing 
the relationship between atmospheric weather variables such as rainfall or wind speed, 
as measured in multiple locations through time, has been 
examined previously in the meteorological literature [e.g., @Serinaldi2008; @li2013]. We therefore 
examined environmental variables from the soil instead, using the Rapid Carbon 
Assessment database [RaCA; @Wills2014].
<!--***DAN: Add Wills and Loecke pending reference if published in time--> 
The database comprises measurements of soil organic carbon and total soil 
nitrogen (megagrams C or N per hectare of soil surface) from `r nrow(d_soilCN)` 
locations across the coterminous United States (Fig. \ref{SM-fig_SoilCNmapUSA} for sampling locations).
Species-trait data were average species basal metabolic rate (BMR, KJ per hour) and 
body mass (grams) for `r nrow(d_birdsBMR)`  species of birds [@Mcnab_2009] 
and `r nrow(d_mammalsBMR)`  species of mammals [@Mcnab_2008].
These data have been much studied, but to our knowledge the copula structure 
of the dependence between metabolic rate and body mass has not been examined. 
Species-trait data such as these reflect the coevolution of the two traits.
Phenological data were aphid first-flight dates from 10 locations (Fig. \ref{SM-fig_aphid_map}) 
across the United Kingdom (UK) for 20 aphid species (Table \ref{SM-tab_aphid_info}), for the 
35 years 1976 to 2010. These time series were computed from the Rothamsted Insect 
Survey (RIS) suction-trap dataset [@Harrington2014; @Bell2015]. The first of our two 
population-level datasets was also derived from the RIS suction-trap data, and comprised 
total counts of aphids trapped for the same locations, species, and years. 
Our second population dataset comprised average annual plankton abundance estimates 
for 14 locations (Fig. \ref{SM-fig_plankton_map})
in the North Sea and British seas for 22 taxa (Table \ref{SM-tab_plankton_info}) for the 
56 years 1958 to 2013. The locations are $2^\circ$ by $2^\circ$ grid cells.
These data were computed from the Continuous Plankton Recorder survey hosted by the 
UK Marine Biological Survey.
Community-level data, obtained from the Cedar Creek Ecosystem Science Reserve, were 
plant aboveground biomass [@ccdataple120] and Shannon's diversity index (computed from plant 
species percent cover data [@ccdatapce120]) for `r nrow(d_cc[[5]])` plots [@ccdataplotse120], 
each 9m by 9m, as sampled in the years 1996-2000 and 2007, each year 
analyzed separately [@Tilman2001; @Tilman2006]. 
Finally, ecosystem functioning data were methane (CH$_{4}$) fluxes between the soil or 
water surface and the troposphere, measured at 13 locations at daily to weekly intervals 
from September 2015 to September 2016 at the Great Miami Wetland mitigation bank in Trotwood, 
Ohio [@Holland1999; @Jarecke2016; @smyth2019using]. Each included location was measured on at least 50 dates. See Appendix \ref{SM-Data_details} for additional data details.

@@

We fit 16 bivariate copula families, exhibiting a variety of 
lower- and upper-tail dependence characteristics, with bivariate datasets. The purpose
of using a large collection of families was solely to include a 
variety of alternative dependence structures to have a robust model
selection and multi-model inference procedure. For that purpose, it 
is not important for the reader to
understand the details of these copulas,
and, additonally, these copulas have been 
described in more detail elsewhere [@Joe1997]; so 
we only briefly identify each family, say a few words about tail dependence 
of the families, and provide pictorial descriptions (Figs 
\ref{figped_N}-\ref{figped_BB1}, \ref{SM-figped_SC} and \ref{SM-figped_G}-\ref{SM-figped_SBB8}). 
The pictorial decriptions can also be used to aid quick comparisons of copula
families when alternative families are being considered for future applications. 
We used several families that can exhibit positive lower-tail dependence 
(of strength depending on parameters) and 
zero upper-tail dependence: the Clayton, survival Gumbel, survival Joe
and survival BB6 families. We used several families that can exhibit positive
upper-tail dependence (strength depends on parameters) and zero lower-tail
dependence: the survival Clayton, Gumbel, Joe and BB6 families. We used
families that show zero upper- and lower-tail dependence: the normal
and Frank families. These families have pdfs symmetric about the line
$v=-u+1$. We also used several families that can show both upper- and
lower-tail dependence, in relative amounts that depend on parameter
values: the BB1, survival BB1, BB7 and survival BB7 families. 
We also used the BB8 family, which shows zero lower-tail dependence,
and zero upper-tail dependence except for a boundary case for the parameters.
And we used the survival BB8 copula, which shows zero upper-tail dependence, 
and zero lower-tail dependence except for a boundary case.
"Survival" families are rotations of the copula with the similar name by
180 degrees. 
See @Joe1997 for details on all these families. We used the implementations
of these families provided in the `VineCopula` package for the R programming
language. See Figs \ref{figped_N}-\ref{figped_BB1}, \ref{SM-figped_SC} and \ref{SM-figped_G}-\ref{SM-figped_SBB8} 
for visual depictions of the pdfs of these copulas and how
pdfs and tail dependence are influenced by the parameters. See 
Table X for a summary of some properties of each family.
<!--***DAN: Shyamolina, please make a table, and replace the X above with 
a citation to that table. Please place the table chunk in such a way that 
this citation becomes table 2, since prior to this point the only main text
table that has been cited is table 1, and we want the table citations to 
come in order. The columns of the table should include: 1) copula family name,
2) number of parameters for the family; 3) info on tail dependence that is
in the above para; 4) a reference to the figure (in the main text or sup mat)
where there is a picture of this family. Anything else? There should be a 
row for each of our 16 copula families.-->
